On Logic Embeddings and G\"odel's God

### 2015-04-04 Papers/2014/WADT/WADT-4pages.tex
improving paper: adding reference, remove typo

\begin{figure}[h]
\begin{flushright}
\begin{minipage}{.97\textwidth}
\VerbatimInput[frame=single,%
               commandchars=\\\{\},%
               fontfamily=courier,
               fontseries=b,%
               fontsize=\scriptsize,%
               %rulecolor=green,%
               %fillcolor=yellow,%
               %formatcom=blue,%
               framerule=1pt,%
               framesep=5pt,
               numbersep=5pt,
               numbers=left]%
               {Quantified_KB.ax}
\end{minipage}
\end{flushright}
                \vskip-.5em
               \caption{\label{fig1}HOL encoding of second-order modal
                 logic KB in THF syntax. 
                 Modal formulas are mapped to HOL 
                 predicates (with type \texttt{\$i>\$o}); 
                 type \texttt{\$i} now stands for possible worlds. 
                 The modal connectives $\neg$ (\texttt{mnot}), $\vee$ (\texttt{mor}) and
                 $\square$ (\texttt{mbox}), universal quantification for
                 individuals (\texttt{mall\_ind}) and for sets of 
                 individuals  (\texttt{mall\_indset}) are introduced in 
                 lines 7-18. 
                 Validity of lifted modal formulas is
                 defined in the standard way (lines 20-21). Symmetry
                 of accessibility relation $r$ is postulated in lines
                 23-26. Hence, second-order KB is realized here; 
                 for logic K the symmetry axiom can be dropped.
                 % \texttt{?}
                   % is the existential quantifier, and
                   % $\neg,\vee,\wedge, and \Rightarrow$ (mat. impl.)
                   % are written as \texttt{\url{~}}, \texttt{|},
                   % \texttt{\&}, and \texttt{=>}. 
                   % Better formatted and easier readable
                   % presentations of our THF code can easily be
                   % generated with the TPTP tools of
                   % Sutcliffe;
                   % here we did minimize space.
                 }
\end{figure}

@@

Ongoing and future work concentrates on the systematic study of
G\"odel's and Scott's proofs. We have also begun to study more recent
variants of the argument
\cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99,fuhrmann05:_exist_notwen,fitting02:_types_tableaus_god,Hajek2002,Hajek2008},
which claim to remedy some fundamental problems of G\"odel's and
Scott's proofs, especially the modal collapse.  One interesting
and very encouraging observation from these studies is, that the
argumentation granularity typical of these philosophy
papers is already within reach of the capabilities 
of our higher-order automated theorem provers.
This provides good evidence for the potential 
relevance of the embedding approach (not only) w.r.t. 
other similar applications in metaphysics.

@@

>>>>

\begin{figure}[t]
\begin{flushright}
\begin{minipage}{.97\textwidth}
\VerbatimInput[frame=single,%
               commandchars=\\\{\},%
               fontfamily=courier,
               fontseries=b,%
               fontsize=\scriptsize,%
               %rulecolor=green,%
               %fillcolor=yellow,%
               %formatcom=blue,%
               framerule=1pt,%
               framesep=5pt,
               numbersep=5pt,
               numbers=left]%
               {Quantified_KB.ax}
\end{minipage}
\end{flushright}
                \vskip-.5em
               \caption{\label{fig1}HOL encoding of second-order modal
                 logic KB in THF syntax. 
                 Modal formulas are mapped to HOL 
                 predicates (with type \texttt{\$i>\$o}); 
                 type \texttt{\$i} now stands for possible worlds. 
                 The modal connectives $\neg$ (\texttt{mnot}), $\vee$ (\texttt{mor}) and
                 $\square$ (\texttt{mbox}), universal quantification for
                 individuals (\texttt{mall\_ind}) and for sets of 
                 individuals  (\texttt{mall\_indset}) are introduced in 
                 lines 7-18. 
                 Validity of lifted modal formulas is
                 defined in the standard way (lines 20-21). Symmetry
                 of accessibility relation $r$ is postulated in lines
                 23-26. Hence, second-order KB is realized here; 
                 for logic K the symmetry axiom can be dropped.
                 % \texttt{?}
                   % is the existential quantifier, and
                   % $\neg,\vee,\wedge, and \Rightarrow$ (mat. impl.)
                   % are written as \texttt{\url{~}}, \texttt{|},
                   % \texttt{\&}, and \texttt{=>}. 
                   % Better formatted and easier readable
                   % presentations of our THF code can easily be
                   % generated with the TPTP tools of
                   % Sutcliffe;
                   % here we did minimize space.
                 }
\end{figure}

@@

Ongoing and future work concentrates on the systematic study of
G\"odel's and Scott's proofs. We have also begun to study more recent
variants of the argument
\cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99,fuhrmann05:_exist_notwen,fitting02:_types_tableaus_god,Hajek2002,Hajek2008},
which claim to remedy some fundamental problems of G\"odel's and
Scott's proofs, especially the modal collapse \cite{C41}.  One interesting
and very encouraging observation from these studies is, that the
argumentation granularity typical of these philosophy
papers is already within reach of the capabilities 
of our higher-order automated theorem provers.
This provides good evidence for the potential 
relevance of the embedding approach (not only) w.r.t. 
other similar applications in metaphysics.

@@

\bibitem{C41}
C. Benzm{\"u}ller and L. Weber and B. Woltzenlogel
Paleo. Computer-Assisted Analysis of the {Anderson-H\'{a}jek}
Ontological Controversy. Handbook of the 1st World Congress on Logic
and Religion, Jo\~ao Pessoa, Brazil, 2015.
